---
title: "Final Project"
author: "Joanna Kennedy"
date: "2022-11-22"
output: github_document
---

knit to html then add to yml


WHEN BUILDING IT ON RSTUDIO NEED TO BUILD WEBSITE BEFOR EYOU PUSH TO GIT HUB

histograms and scatterplots - min plot number
dendrogram

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(stringr)
library(lubridate)
library(tidyverse)
library(ggrepel)
library(caret)
library(MLmetrics)
library(stats)
library(dplyr)
library(glmnet)
```


```{r}
setwd("C:/Users/jjken/Desktop/Harvard/Fall 2022/BST 260/finalprojectwebsite")
df = read.csv(file = 'Student Mental health.csv')
```

Cleaning
```{r}
head(df)
```

```{r}
apply(df, 2, unique)
```


???????WHEN NUMBER OF YES IS SMALL SHOULD I SPLIT THE TRAIN AND TEST TO INCLUDE ALL YES IN BOTH TRAIN AND TEST (or some for even amount) - should I double dip
--> should I decrease the size of testing to get half and half and have all in training (OR 1/3)
```{r}

for (i in 1:ncol(df)){
  
  temp <- as.data.frame(df[,i])
  nm <- colnames(df)[i]
  colnames(temp) <- "response"
  
  counts <- temp %>% group_by(response) %>% dplyr::summarise(count = n())
  
  plot<- ggplot(data=counts, aes(x=response, y=count, fill=response)) +
    geom_bar(stat="identity") +
    geom_text(aes(label=count)) +
    theme(axis.text.x = element_text(size = 6, angle = 90)) +
    ggtitle(nm)
  print(plot)
  
}
```


```{r}
colnames(df) = c("day","gender","age", "course","grade","gpa","married", "depression","anxiety","panic_attack", "treatment")
```

```{r}
sapply(df, function(x) sum(is.na(x)))
```

```{r}
df <- df %>% filter(!is.na(df$age))
```

```{r}
df <- as.data.frame(apply(df, 2, tolower))
```


year and month are the same so just day and time are different
```{r}
df <- separate(df, col=day, into=c("day", "time"), sep = " ")
df <- separate(df, col=day, into=c("day", "rest"), sep = "/")
df$rest <- NULL
df <- separate(df, col=time, into=c("hour", "minute"), sep = ":")
```

round hour and get rid of minute
```{r}
df$hour <- as.numeric(df$hour)
df$minute <- as.numeric(df$minute)
df$round <- 0
df[which(df$minute >29), 'round'] <- 1
df$hour <- df$hour + df$round
df$round <- NULL
df$minute <- NULL
```

```{r}
df$course <- sapply(df$course, str_trim)
df$course <- df$course %>% str_replace("^engin.*","engineering") %>% str_replace("^fiqh.*","fiqh") %>% str_replace("^law.*", "laws") %>% str_replace(".*irkhs$", "kirkhs") %>% str_replace(".*nursing$", "nursing")
df %>% arrange(course) %>% group_by(course) %>% dplyr::summarise(count = n()) %>% arrange(-count)
```

```{r}
df$grade <- df$grade %>% str_replace("year 1","9") %>% str_replace("year 2","10") %>% str_replace("year 3", "11") %>% str_replace("year 4", "12")
```

```{r}
df$gpa <- sapply(df$gpa, str_trim)
df <- separate(df, col=gpa, into=c("low", "gpa"), sep = " - ")
df$low <- NULL
```

```{r}

for (i in 1:ncol(df)){
  
  temp <- as.data.frame(df[,i])
  nm <- colnames(df)[i]
  colnames(temp) <- "response"
  
  counts <- as.data.frame(temp) %>% group_by(response) %>% dplyr::summarise(count = n())
  
  plot<- ggplot(data=counts, aes(x=response, y=count, fill=response)) +
    geom_bar(stat="identity") +
    geom_text(aes(label=count)) +
    theme(axis.text.x = element_text(size = 6, angle = 90)) +
    ggtitle(nm)
  print(plot)
  
}

#ggplot(age_counts_mental_health, aes(x=mental_health, y=count)) +
#  geom_point() +
  

```


```{r}
df[,c(1,2,4,6,7)]<-as.data.frame(apply(df[,c(1,2,4,6,7)], 2, as.numeric))
numerical <- df[,c(1,2,4,6,7)]
numerical_colnames <- c("day","hour", "age","grade","gpa")
categorical <- df[,c(5)]
categorical_colnames <- c("course")
binary <- df[,c(3,8:12)]
binary_colnames <- c("gender", "married", "depression", "anxiety", "panic_attack", "treatment") #"engineering", "bcs")
```

```{r}
#apply(numerical,2,class)
#apply(numerical,2,hist)
```
```{r}
numerical
```


```{r}
plot(cbind(numerical,binary))
plot(numerical)
```

```{r}

test <- df %>% group_by(grade,course) %>% dplyr::summarise(group_gpa=mean(gpa), count = n()) %>% filter(count>2)

ggplot(test, aes(x=grade,y=group_gpa, color=course)) +
  geom_line() +
  geom_point()

```

```{r}
test <- df %>% group_by(grade,course) %>% dplyr::summarise(group_gpa=mean(gpa), count = n()) %>% filter(count>2)

ggplot(test, aes(grade, group_gpa)) + 
  geom_jitter(aes(col=course)) + 
  geom_smooth(aes(col=course), method="lm", se=F)
```


```{r}

explore <- df

label_mental_health <- function(df,last_column){
  df$mental_health <- ""
  for (participant in 1:nrow(df)){
    for (condition in 9:last_column){
      if (df[participant,][condition] == "yes"){
        df[participant,][last_column+1] <- paste(df[participant,][last_column+1],colnames(df[condition]))
      }
    }
  }
  df$mental_health <- sapply(df$mental_health, str_squish)
  return (df)
}

explore <- label_mental_health(explore, dim(explore)[2])
```


```{r}
explore %>% group_by(mental_health) %>% dplyr::summarise(count = n())
```
```{r}

for(i in 1:(ncol(numerical))){
  
  temp <- as.data.frame(numerical[,i])
  nm <- colnames(numerical)[i]
  colnames(temp) <- "response"
  
  temp$mental_health <- explore$mental_health
  
  counts <- as.data.frame(temp) %>% group_by(response,mental_health) %>% dplyr::summarise(count = n())
  
  plot<- ggplot(data=counts, aes(x=response, y=count, fill=response)) +
    geom_bar(stat="identity") +
    geom_text(aes(label=count)) +
    facet_wrap(~mental_health, ncol=3) +
    theme(axis.text.x = element_text(size = 6, angle = 90)) +
    ggtitle(nm)
  print(plot)
  
}

```


Chi Squared Test

```{r}

two_by_two <- data.frame(
  depression = c("no", "yes"), 
  men = 
    c(sum(df$gender == "male" & df$depression == "no"), 
      sum(df$gender == "male" & df$depression == "yes")), 
  women = 
    c(sum(df$gender == "female" & df$depression == "no"), 
      sum(df$gender == "female" & df$depression == "yes")))

two_by_two

chisq_test <- two_by_two |> select(-depression) |> chisq.test()
chisq_test

```

```{r}
length(binary_colnames)
length(binary)
```
```{r}
binary_colnames
```


```{r}
ncol(binary)
binary
```


```{r}
association <- c()
for(i in 1:(ncol(binary)-1)){
  for(j in (i+1):(ncol(binary))){
    if (i != j){
      print(binary_colnames[i])
      print(binary_colnames[j])
      test <- chisq.test(df[,i], df[,j])
      print(test)
      if (test$p.value < 0.05) {
        association <- c(association, paste(binary_colnames[i], binary_colnames[j], sep = " vs. "))
      }
    }
    
  }
}
```
```{r}
association
```

Regression

```{r}

non_normal_numerical = c()
for (i in 1:ncol(numerical)){
  test <- shapiro.test(numerical[,i])
  print(test)
  
  if (test$p.value < 0.05)
    non_normal_numerical <- c(non_normal_numerical, numerical_colnames[i])
}

non_normal_numerical

```

```{r}

for(i in 1:(ncol(numerical)-1)) {
  for(j in (i+1):ncol(numerical)){
    for(k in binary_colnames){
      plot <- ggplot(df,aes(x=df[,numerical_colnames[i]], y=df[,numerical_colnames[j]], color=df[,k], group = df[,k])) +
        #geom_boxplot() +
        geom_point(position=position_jitter(h=0.07,w=0.07)) +
        ylab(numerical_colnames[j]) +
        xlab(numerical_colnames[i]) +
        labs(color = k)
      print(plot)
    }
  }
}


```

I acknowledge gender as a categorical value rather than a binary value, but within the context of this datasets, students were categorized as either female or male.

```{r}
df
```


```{r}
head(train)
```


```{r}
#c("gender", "married", "depression", "anxiety", "panic_attack", "treatment", "engineering")

traini <- sample(nrow(df), round(0.8*nrow(df)))
testi <- c(1:nrow(df))[-traini]
train <- df[traini, ]
test <- df[testi, ]


train$gender <- ifelse(train$gender == 'female',1,0)
train$married <- ifelse(train$married == 'yes',1,0)
train$depression <- ifelse(train$depression == 'yes',1,0)
train$anxiety <- ifelse(train$anxiety == 'yes',1,0)
train$panic_attack <- ifelse(train$panic_attack == 'yes',1,0)
train$treatment <- ifelse(train$treatment == 'yes',1,0)
train$engineering <- ifelse(train$course == 'engineering',1,0)
train$bcs <- ifelse(train$course == 'bcs',1,0)

test$gender <- ifelse(test$gender == 'female',1,0)
test$married <- ifelse(test$married == 'yes',1,0)
test$depression <- ifelse(test$depression == 'yes',1,0)
test$anxiety <- ifelse(test$anxiety == 'yes',1,0)
test$panic_attack <- ifelse(test$panic_attack == 'yes',1,0)
test$treatment <- ifelse(test$treatment == 'yes',1,0)
test$engineering <- ifelse(test$course == 'engineering',1,0)
test$bcs <- ifelse(test$course == 'bcs',1,0)

train <- train[,-5]
test <- test[,-5]

train <- train[,c(1,2,4:6,3,7:ncol(train))]
test <- test[,c(1,2,4:6,3,7:ncol(test))]

train
test

```



Lasso Regression

source: https://www.pluralsight.com/guides/linear-lasso-and-ridge-regression-with-r

```{r}
eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  
  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square
)
  
}
```


```{r}

dummies_train <- dummyVars(depression~ ., data = train)
dummies_test <- dummyVars(depression~ ., data = test)
train_dummies = predict(dummies_train, newdata = train)
test_dummies = predict(dummies_test, newdata = test)

x = as.matrix(train_dummies)
y_train = train$depression

x_test = as.matrix(test_dummies)
y_test = test$depression

```


```{r}

lambdas <- 10^seq(2, -3, by = -.1)

# Setting alpha = 1 implements lasso regression
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

# Best 
lambda_best <- lasso_reg$lambda.min 
lambda_best

```

```{r}
plot(lasso_reg)
```

want smaller RMSE
```{r}

lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y_train, predictions_train, train)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test, test)

```

```{r}
lasso_model$beta
```



```{r}
colnames(train)
```


GLM AND KNN

k
accuracy
binary

```{r}
for (k in 1:1){
  knn_fit <- knn3(depression~., data=train, k=k) 
  p_hat_knn_train <- predict(knn_fit, train, type="prob")
  y_hat_knn_train <- factor(ifelse(p_hat_knn_train[,2]>0.5,1,0))
  confusionMatrix(y_hat_knn_train, factor(train$depression))
  #print(k)
  #print(confusionMatrix(y_hat_knn_train, factor(train$depression))$overall["Accuracy"])
  
  
  p_hat_knn_test <- predict(knn_fit, test, type="prob")
  y_hat_knn_test <- factor(ifelse(p_hat_knn_test[,2]>0.5,1,0))
  confusionMatrix(y_hat_knn_test, factor(test$depression))
  #print(confusionMatrix(y_hat_knn_test, factor(test$depression))$overall["Accuracy"])
}



for (k in 1:10){
  knn_fit <- knn3(anxiety~., data=train, k=k) 
  p_hat_knn_train <- predict(knn_fit, train, type="prob")
  y_hat_knn_train <- factor(ifelse(p_hat_knn_train[,2]>0.5,1,0))
  confusionMatrix(y_hat_knn_train, factor(train$anxiety))
  #print(k)
  #print(confusionMatrix(y_hat_knn_train, factor(train$anxiety))$overall["Accuracy"])
  
  
  p_hat_knn_test <- predict(knn_fit, test, type="prob")
  y_hat_knn_test <- factor(ifelse(p_hat_knn_test[,2]>0.5,1,0))
  confusionMatrix(y_hat_knn_test, factor(test$anxiety))
  #print(confusionMatrix(y_hat_knn_test, factor(test$anxiety))$overall["Accuracy"])
}
```


??????????????????????WARNING
```{r}
df
```


```{r}
train
```


```{r}
test
```

```{r}

```

```{r}
head(train)
```

```{r}

ks = c()
train_accuracy = c()
test_accuracy = c()
binary_labels = c()

for (i in 6:ncol(train)){
  
  colname <- colnames(train)[i]
  if (colname == "treatment"){
    next
  }
  
  colnames(train)[i] <- "label"
  colnames(test)[i] <- "label"
  
  print(colname)
  
  glm_1 <- glm(label~., data=train, family="binomial")
  #predict(glm_1, train)
  plot(glm_1)
  
  for (k in 1:10){
    
    knn_fit <- knn3(label~., data=train, k=k) 
    p_hat_knn_train <- predict(knn_fit, train, type="prob")
    y_hat_knn_train <- factor(ifelse(p_hat_knn_train[,2]>0.5,1,0))
    confusionMatrix(y_hat_knn_train, factor(train$label))
    #print(confusionMatrix(y_hat_knn_train, factor(train$label))$overall["Accuracy"])
    
    p_hat_knn_test <- predict(knn_fit, test, type="prob")
    y_hat_knn_test <- factor(ifelse(p_hat_knn_test[,2]>0.5,1,0))
    confusionMatrix(y_hat_knn_test, factor(test$label))
    #print(confusionMatrix(y_hat_knn_test, factor(test$label))$overall["Accuracy"])
    
    ks <- c(ks, k)
    train_accuracy <- c(train_accuracy, confusionMatrix(y_hat_knn_train, factor(train$label))$overall["Accuracy"])
    test_accuracy <- c(test_accuracy, confusionMatrix(y_hat_knn_test, factor(test$label))$overall["Accuracy"])
    binary_labels <- c(binary_labels, colname)
  
  }
  
  colnames(train)[i] <- colname
  colnames(test)[i] <- colname
}


```



```{r}

binary_labels
ks
train_accuracy
test_accuracy
knn_dataframe <- data.frame(ks,train_accuracy,test_accuracy,binary_labels)
knn_dataframe

```

```{r}
ggplot(knn_dataframe, aes(x=ks)) +
  geom_line(aes(y = train_accuracy), color = "red") + 
  geom_line(aes(y = test_accuracy), color = "blue") +
  facet_wrap(~binary_labels)
```








